Reconstruction of Large Scale Structure by Nonlinear Displacement Fields

### 2016-09-20 ms_disp1/ms_disp1.tex
haoran: bug fix in initial_conditions.f90; disp1 paper update

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}

@@

\subsection{Simulation}
We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters are in accordance with Planck. Initial conditions are generated at redshift $z=100$ using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho$ at late stages.

Initial conditions, get linear density field $\delta_L$. Run simulation by code {\tt CUBE} and get the nonlinear density field $\delta$ at $z=0$. {\tt CUBE} uses particle-ID (PID) to record the initial location of particles, and the information is tracked until the $z=0$ and we can get the displacement vector ${\bs s}\equiv{\bs x}-{\bs q}$ for every particle. Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs s}({\bs q})$.

\subsection{Reconstruction}
Get the raw reconstructed density field $\delta_R=-\nabla\cdot{\bs s}({\bs q})$.

@@

Wiener filter.
Define
\begin{equation}
    r\equiv \frac{\left\langle\delta_L\delta_R\right\rangle}
    {\sqrt{\left\langle\delta_L\delta_L\right\rangle
    \left\langle\delta_R\delta_R\right\rangle}}
\end{equation}
and
\begin{equation}
    b^2\equiv \frac{\left\langle\delta_R\delta_R\right\rangle}
    {\left\langle\delta_L\delta_L\right\rangle}.
\end{equation}

We try to decompose $\delta_L$ into two components,
\begin{equation}\label{eq.decompose}
    \delta_L=r'\delta_R+\delta_N,
\end{equation}
where $r'\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution. To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
\begin{equation}\label{eq.correlate}
    \left\langle\delta_L\delta_R\right\rangle
    =r'\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_R\right\rangle,
\end{equation}
and vanish the second term in the right-hand-side. Then
\begin{equation}
    r'=\frac{\left\langle\delta_L\delta_R\right\rangle}
    {\left\langle\delta_R\delta_R\right\rangle}=\frac{r}{b}.
\end{equation}
Then the mode-coupling term 
\begin{equation}
    \delta_N=\delta_L-\frac{r}{b}\delta_R.
\end{equation}
We next square equation (\ref{eq.decompose}) and gives
\begin{equation}\label{eq.square}
    \left\langle\delta_L\delta_L\right\rangle
    =\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_N\right\rangle
    +2\frac{r}{b}\left\langle\delta_R\delta_N\right\rangle,
\end{equation}
and the third term of right-hand-side also vanishes. As a result,
\begin{equation}
    \left\langle\delta_N\delta_N\right\rangle
    =\frac{1-r^2}{b^2}
    \left\langle\delta_R\delta_R\right\rangle
\end{equation}
and a Wiener filter can be constructed as
\begin{equation}
    W=\frac{r^2}{r^2+1-r^2}=r^2.
\end{equation}

From equation (\ref{eq.decompose}), multiplying $W$, we obtain the optimal reconstructed linear density field by
\begin{equation}
    \tilde\delta_L=W\frac{r}{b}\delta_R=\frac{r^3}{b}\delta_R.
\end{equation}
From equation (\ref{eq.square}), multiplying $W^2$, the reconstructed linear power spectrum is giving by
\begin{equation}
\begin{aligned}
    \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
    &=r^2\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle +r^2\left\langle\delta_N\delta_N\right\rangle  \\
    &=\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle.
\end{aligned}
\end{equation}

>>>>

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}
\newcommand{\lb}{\left\langle}
\newcommand{\rb}{\right\rangle}

@@

\subsection{Simulation}
We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters are in accordance with Planck. Initial conditions are generated at redshift $z=50$ using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates ${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ to describe the density fluctuations. The primordial linear density field $\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth factor.

\subsection{Reconstruction}
In the simulation, we uses particle-ID (PID) to record the initial location ${\bs q}$ of particles, and the information is tracked until the $z=0$ and we can get the Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every particle. Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}

To quantify the linear information in the reconstructed density field, we decompose $\delta_R$ in Fourier space as
\begin{equation}\label{eq.decompose}
    \delta_R(k)=r'\delta_L+\delta_N,
\end{equation}
where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_L$ gives
\begin{equation}
    P_{LR}=r'P_{LL}+P_{LN}
\end{equation}
where $P_{ij}\equiv\lb\delta_i\delta_j\rb$ denotes the cross power spectrum. Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$. With the definition of cross correlation coefficient $r\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
\begin{equation}
    P_{RR}=r^2b^2P_{LL}+P_{NN},
\end{equation}
and $P_{NN}=(1-r^2)P_{RR}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
\begin{equation}
    W(k)=\frac{r^2b^2P_{LL}}{r^2b^2P_{LL}+P_{NN}}=r^2.
\end{equation}
The estimated linear density, or the optimal reconstructed density is given by
\begin{equation}
    \tilde\delta_R=Wb^{-1}\delta_R,
\end{equation}
and the optimal reconstructed power spectrum is given by
\begin{equation}
    \tilde P=W^2b^{-2}P_{RR}=W^2P_{LL}+W^2b^{-2}P_{NN}.
\end{equation}

@@
